---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rtlr

<!-- badges: start -->
<!-- badges: end -->

The goal of rtlr is to help you print correctly right-to-left text in the console and in plots.

## Installation

You can install the development version of rtlr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("matanhakim/rtlr")
```

## Example

This is a basic example of fixing an RTL issue when writing in Hebrew. Hebrew is written right-to-left, so punctuation marks (such as "," or ".") should come at the left side, i.e. the end, of a sentance.

```{r example}
library(rtlr)
x <- "הנקודה צריכה להיות בסוף המשפט."
# `x` says "the dot should come at the end of the sentance." in Hebrew.

cat(x)
cat(str_rtl(x))
```
Notice how the dot ends up in the left side, where it should be.


You can also use it in plots, for example:

```{r plot_1, out.width="50%"}
library(ggplot2)
theme_update(text = element_text(size = 20))
ggplot(mtcars, aes(cyl)) +
  geom_bar() +
  xlab("اسطوانات!")
# `xlab` says "cylinders!" in Arabic.
```

Notice how the exclamation mark is on the right-hand side, but it should be on the left, as Arabic is another right-to-left language.
use `str_rtl()` to solve it, like this:

```{r plot_2, out.width="50%"}
ggplot(mtcars, aes(cyl)) +
  geom_bar() +
  xlab(str_rtl("اسطوانات!"))
```

Now the exclamation mark is on the left-hand side, like it should.

Additionally, sometimes you wish to break up lines within a plot. The easiest way to do it is by pasting `\n`, but this results yet again with a problem in the location of non-RTL characters. Notice how the parentheses on the top `xlab()` row are correctly placed on the left side, but the parantheses on the bottom row are wrongly placed on the right side.
```{r plot_3, out.width="50%"}
ggplot(mtcars, aes(cyl)) +
  geom_bar() +
  xlab(paste0("اسطوانات! (عدد المكابس)", "\n", "عام (2023)"))
# `xlab()` says in Arabic:
# "cylinders! (The number of pistons)"
# "Year (2023)"
```

This can be easily solved with `str_rtl()` and `multiline = TRUE`:
```{r plot_4, out.width="50%"}
ggplot(mtcars, aes(cyl)) +
  geom_bar() +
  xlab(str_rtl("اسطوانات! (عدد المكابس)", "عام (2023)", multiline = TRUE))
# `xlab()` says in Arabic:
# "cylinders! (The number of pistons)"
# "Year (2023)"
```
